Adaptive Approximation for Multivariate Linear Problems \\ with Inputs Lying in a Cone

### 2018-11-06 RICAM2018NovTalk.tex
revising

    \item \hyperlink{BallsWontHelp}{\beamergotobutton{``But I really like balls!''}}
\end{itemize}

@@

\begin{align*}
    \cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \bignorm[1,\vlambda]{\hf} : = \norm[1]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractabililty}\end{minipage}\\
    u_{\vj} &= \text{Cosine/Sine or Walsh} \qquad S(u_{\vj}) = \int_{[0,1]^d} u_\vj (\vx) \, \dif \vx = \delta_{\vj,\vzero}, \qquad \cg : = \reals \\
    &\begin{array}{l}\Sapp(f,n) = \frac 1n \sum_{i=1}^n f(\vx_i)  \\[4ex]
    \abs{S(f) - \Sapp(f,n)} \le \sum_{\vzero \ne \vj \in \text{dual set}} \bigabs{\hf(\vj)}
    \end{array}
    \quad \raisebox{-8ex}{\includegraphics[width=3cm]{ProgramsImages/ShiftedLatticePoints.eps} \qquad
    \includegraphics[width=3cm]{ProgramsImages/SSobolPoints.eps}}
\end{align*}

@@

\begin{frame}{Cones of Integrands Whose Fourier Coefficients Decay Steadily\footfullcite{HicJim16a,JimHic16a}}

@@

\begin{align*}
n_0 & = 0, \quad n_k = 2^{k-1}, \ k \in \naturals, \qquad \vj \text{ ordered}, \quad \text{true coef.} = \tf(\vj) \approx \tf_{\text{disc}}(\vj) = \text{discrete coef.}\\
	 \sigma_k(f) & = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad  \hsigma_{k,\ell}(f)  = \sum_{i=n_{k-1}+1}^{n_k} \sum_{m=1}^{\infty} \bigabs{\hf(\vj_{i+mn_{\ell}})}, \quad \sigma_{\textup{disc},k}(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf_{\text{disc}}(\vj_i)} \\
	 \cc &: = \left\{f \in \cf :  
	 \sigma_\ell(f) \le a_1 b_1^{\ell - k}\sigma_k(f), \   \hsigma_{k,\ell}(f) \le a_2 b_2^{\ell - k} \sigma_\ell(f) \quad \forall k \le \ell \right\} \quad a_1, a_2 > 1 > b_1, b_2\\
      \MoveEqLeft{\Sapp(f,n_k) = \frac 1{n_k} \sum_{i=1}^{n_k} f(\vx_i)  \quad
    \abs{S(f) - \Sapp(f,n_k)} \le \sum_{\vzero \ne \vj \in \text{dual set}} \bigabs{\hf(\vj)} \le C(k) \sigma_{\textup{disc},k}(f)} \\
    \alert{A(f,\varepsilon)} & \alert{= S(f,n_k) \text{ for the smallest $k$ satisfying } C(k) \sigma_{\textup{disc},k}(f) \le \varepsilon}
\end{align*}

\vspace{-4ex}
\alert{Have } $\cost(f,A,\varepsilon)$; \quad \alert{No } $\cost(A,\cc,\varepsilon,\rho)$,  $\comp(\ca(\cc),\varepsilon,\rho)$, or  tractability results \alert{yet}

@@

\begin{align*}
    \cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \bignorm[2,\vlambda]{\hf} : = \norm[2]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[2]{\hg}\biggr \}, \qquad S(u_{\vj}) = v_{\vj} \\[-2ex]
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      n_0 < n_1 < n_2 < \cdots,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n) &= \sum_{i=1}^n \hf(\vj_i) v_{\vj_i} \text{ is a optimal for fixed }n, \qquad \norm[\cg]{S(f) - \Sapp(f,n)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= S(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
\end{align*}

>>>>

\end{itemize}

@@

\begin{align*}
    f &= \sum_{\vj \in \cn} \hf(\vj) u_{\vj}, \qquad
    u_{\vj} = \text{Cosine/Sine or Walsh}, \qquad S(u_{\vj}) = \int_{[0,1]^d} u_\vj (\vx) \, \dif \vx = \delta_{\vj,\vzero}\\
    &\begin{array}{l}\Sapp(f,n) = \frac 1n \sum_{i=1}^n f(\vx_i)  \\[4ex]
    \abs{S(f) - \Sapp(f,n)} \le \sum_{\vzero \ne \vj \in \text{dual set}} \bigabs{\hf(\vj)}
    \end{array}
    \quad \raisebox{-8ex}{\includegraphics[width=3cm]{ProgramsImages/ShiftedLatticePoints.eps} \qquad
    \includegraphics[width=3cm]{ProgramsImages/SSobolPoints.eps}}
\end{align*}

@@

\begin{frame}{Cones of Integrands Whose Fourier Series Coefficients Decay Steadily\footfullcite{HicJim16a,JimHic16a}}

@@

\begin{align*}
n_0 & = 0, \quad n_k = 2^{k-1}, \ k \in \naturals, \qquad \vj \text{ ordered}, \quad \text{true coef.} = \hf(\vj) \approx \tf_{\text{disc}}(\vj) = \text{discrete coef.}\\
	 \sigma_k(f) & = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad  \hsigma_{k,\ell}(f)  = \sum_{i=n_{k-1}+1}^{n_k} \sum_{m=1}^{\infty} \bigabs{\hf(\vj_{i+mn_{\ell}})}, \quad \sigma_{\textup{disc},k}(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf_{\text{disc}}(\vj_i)} \\
	 \cc &: = \left\{f \in \cf :  
	 \sigma_\ell(f) \le a_1 b_1^{\ell - k}\sigma_k(f), \   \hsigma_{k,\ell}(f) \le a_2 b_2^{\ell - k} \sigma_\ell(f) \quad \forall k \le \ell \right\} \quad a_1, a_2 > 1 > b_1, b_2\\
      \MoveEqLeft{\Sapp(f,n_k) = \frac 1{n_k} \sum_{i=1}^{n_k} f(\vx_i)  \quad
    \abs{S(f) - \Sapp(f,n_k)} \le \sum_{\vzero \ne \vj \in \text{dual set}} \bigabs{\hf(\vj)} \le C(k) \sigma_{\textup{disc},k}(f)} \\
    \alert{A(f,\varepsilon)} & \alert{= S(f,n_k) \text{ for the smallest $k$ satisfying } C(k) \sigma_{\textup{disc},k}(f) \le \varepsilon}
\end{align*}

\vspace{-4ex}
\alert{Have} $\cost(f,A,\varepsilon)$; \quad \alert{No} $\cost(A,\cc,\varepsilon,\rho)$,  $\comp(\ca(\cc),\varepsilon,\rho)$, or  tractability results \alert{yet}

@@

\begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \norm[2]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[2]{\hg}\biggr \}, \qquad v_{\vj} = S(u_{\vj}) \\[-2ex]}
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      n_0 < n_1 < n_2 < \cdots,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n_k) &= \sum_{i=1}^{n_k} \hf(\vj_i) v_{\vj_i} \text{ is optimal for fixed }n_k, \qquad \norm[\cg]{S(f) - \Sapp(f,n_k)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
 	 \only<2>{\\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}}} \\
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) \text{ essentially no worse than } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}}}
\end{align*}

\begin{frame}{\emph{Function Approximation} when Function Values Are Expensive}

    \begin{align*}
    \only<1>{\cf &:= \left \{ f = \sum_{\vj \in \natzero^d} \hf(\vj) u_{\vj} : \norm[\cf]{f} :=  \norm[\infty]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \qquad \begin{minipage}{3cm}\raggedright\alert{$\vlambda$ affects convergence rate \& tractability}\end{minipage}\\
    \cg &: = \biggl \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[1]{\hg}\biggr \}, \qquad v_{\vj} = S(u_{\vj}) = u_{\vj} \\[-2ex]}
    \lambda_{\vj} & = \Gamma_{\norm[0]{\vj}} \prod_{\substack{\ell = 1 \\ j_{\ell} > 0}}^d \gamma_\ell s_{j_\ell} \quad \text{\alert{product, order, and smoothness dependent weights}} \\
      & \alert{\lambda_{\vj_1} \ge \lambda_{\vj_2} \ge \cdots}, \qquad
      \Sapp(f,n) = \sum_{i=1}^{n} \hf(\vj_i) u_{\vj_i} ,  \qquad
 	\sigma_k(f) = \sum_{i=n_{k-1}+1}^{n_k}  \bigabs{\hf(\vj_i)}, \quad k \in \naturals \\
     \cc &: = \left\{f \in \cf :  
 	 \sigma_\ell(f) \le a b^{\ell - k}\sigma_k(f) \quad \forall k \le \ell \right\} \quad a > 1 > b \quad \text{\alert{series coef.\ decay steadily}}\\
 	 \Sapp(f,n_k) &= \sum_{i=1}^{n_k} \hf(\vj_i) v_{\vj_i} \text{ is optimal for fixed }n_k, \qquad \norm[\cg]{S(f) - \Sapp(f,n_k)} \le \frac{ab \sigma_k(f)}{\sqrt{1 - b^2}} \\[-1ex]
 	 \alert{A(f,\varepsilon)} & \alert{= \Sapp(f,n_k) \text{ for the smallest $k$ satisfying } \sigma_k(f) \le \frac{\varepsilon \sqrt{1 - b^2}}{ab}}
 	 \only<2>{\\ 
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) = n_{\ell^\dagger}, \quad 
 	 \ell^\dagger \le \min \left \{\ell \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{\ell-1} \frac{b^{2(k-\ell)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{\ell-1}+1}^2}\right]   \right\}}} \\
 	 \MoveEqLeft{\alert{\cost(A,\cc,\varepsilon,\rho) \text{ essentially no worse than } \comp(\ca(\cc,\LambdaAll),\varepsilon,\rho)}}}
\end{align*}
